Slavnov products and KP tau functions

### 2025-04-18 draft.tex
calculate the BA in the z-coordinate system

Using these expressions, we can write 
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \det_{j,k}\left[
  \oint_{\gamma_k} \frac{dw}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{- j -1} ( \lambda w - 1) \mathcal{Y}_k(w)}{\lambda(w - v_k)} 
  \right]\; .
\end{equation}
that in terms of the \(z\)-coordinates becomes
\begin{equation}
\begin{split}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v})
  & = \det_{j,k}\left[\oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{( \lambda w - 1) w^{n-j-1}\mathcal{Y}_k(w)}{\lambda(w - v_k)} \right]\\
  & = \det_{j,k}\left[
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{w^{n-j-1}\widetilde{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    \right]
\end{split}
\end{equation}
where we have defined the function 
\begin{equation}
  \widetilde{\mathcal{Y}}_k(w; \lambda)
  = \frac{( \lambda w - 1)}{\lambda} \mathcal{Y}_k(w) \; ,
\end{equation}
From these expressions, it is easy to see that for \(j=n\), then \(w =
0\) is a pole and we must pick this term in the integration curve
\(\gamma_k\).

Let us define the components
\begin{equation}
  \begin{split}
    \widetilde{\mathcal{K}}_{kj} = &
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}
    \frac{w^{n-j-1}\widetilde{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    + \delta_{nj} F_k(\lambda) \\
    & = \sum_{\ell = 1}^n  \widetilde{\Omega}_{k \ell} \left( \frac{z_\ell^{n-j}}{\prod_{s\neq \ell} (z_\ell - z_s)} \right)
    + F_k(\lambda)\delta_{nj}  \\
  \end{split}
\end{equation}
where
\begin{equation}
  \widetilde{\Omega}_{k\ell}
  = \frac{1}{z_\ell} \frac{\widetilde{\mathcal{Y}}_k(\lambda, z_\ell)}{z_\ell - v_k}
  = \frac{(\lambda z_\ell -1)}{\lambda z_\ell} {\Omega}_{k\ell} 
\; , \qquad 
 F_k(\lambda) = (-1)^n \frac{\widetilde{\mathcal{Y}}_k(\lambda; 0)}{v_k }  \prod_{s=1}^n z_s^{-1} \; .
\end{equation}

In order to simplify our analysis, let us define one more matrix
\(\widehat{\mathcal{K}}\) by its components
\begin{equation}
    \widehat{\mathcal{K}}_{kj} = 
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}
    \frac{w^{n-j-1}\widetilde{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}\; .
\end{equation}
We conclude that \(\widetilde{\mathcal{K}}\) can be obtained from 
\(\widehat{\mathcal{K}}\) by adding to each component of the \(n\)-th row
the corresponding component of the vector \(\bm{F} = (F_1, \dots , F_n)\). 
Let us write this operation as 
\begin{equation}
  \widetilde{\mathcal{K}} = \widehat{\mathcal{K}} + \bm{F}_{[\rightarrow n-row]}\; ,
\end{equation}

>>>>

Using these expressions, we can write 
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \det_{j,k}\left[
  \oint_{\gamma_k} \frac{dw}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{- j -1} ( \lambda w - 1) \mathcal{Y}_k(w)}{\lambda(w - v_k)} 
  \right]\; .
\end{equation}
that in terms of the \(z\)-coordinates becomes
\begin{equation}
\begin{split}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v})
  & = \det_{j,k}\left[\oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{( \lambda w - 1) w^{n-j-1}\mathcal{Y}_k(w)}{\lambda(w - v_k)} \right]\\
  & = \det_{j,k}\left[
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    \right]
\end{split}
\end{equation}
where we have defined the function 
\begin{equation}
  \widehat{\mathcal{Y}}_k(w; \lambda)
  = \frac{( \lambda w - 1)}{\lambda} \mathcal{Y}_k(w) \; ,
\end{equation}
From these expressions, it is easy to see that for \(j=n\), then \(w =
0\) is a pole and we must pick this term in the integration curve
\(\gamma_k\).

Let us define the components
\begin{equation}
  \begin{split}
    \widetilde{\mathcal{K}}_{kj} = &
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    + \delta_{nj} F_k(\lambda) \\
    & = \sum_{\ell = 1}^n  \widehat{\Omega}_{k \ell} \left( \frac{z_\ell^{n-j}}{\prod_{s\neq \ell} (z_\ell - z_s)} \right)
    + F_k(\lambda)\delta_{nj}  \\
  \end{split}
\end{equation}
where
\begin{equation}
  \widehat{\Omega}_{k\ell}
  = \frac{1}{z_\ell} \frac{\widetilde{\mathcal{Y}}_k(\lambda, z_\ell)}{z_\ell - v_k}
  = \frac{(\lambda z_\ell -1)}{\lambda z_\ell} {\Omega}_{k\ell} 
\; , \qquad 
 F_k(\lambda) = (-1)^n \frac{\widehat{\mathcal{Y}}_k(\lambda; 0)}{v_k }  \prod_{s=1}^n z_s^{-1} \; .
\end{equation}

In order to simplify our analysis, let us define one more matrix
\(\widehat{\mathcal{K}}\) by its components
\begin{equation}
    \widehat{\mathcal{K}}_{kj} = 
    \sum_{\ell = 1}^n  \widehat{\Omega}_{k \ell}
    \left( \frac{z_\ell^{n-j}}{\prod_{s\neq \ell} (z_\ell - z_s)} \right)
    \; .
\end{equation}
We conclude that \(\widetilde{\mathcal{K}}\) can be obtained from 
\(\widehat{\mathcal{K}}\) by adding to each component of the \(n\)-th row
the corresponding component of the vector \(\bm{F} = (F_1, \dots , F_n)\). 
Let us write this operation as 
\begin{equation}
  \widetilde{\mathcal{K}} = \widehat{\mathcal{K}} + \bm{0}_{[ \bm{F} \to n-row]}\; . 
\end{equation}
Where we use the notation \(\mathcal{A}_{[ \bm{F} \to j-row]} \) to denote
the replacement of the \(j-th\) row of the matrix with the vector \(\bm{F}\).
Above we have considered this operation with the zero matrix \(\bm{0}\).

By the multilinearity of the determinant, we can write
\begin{equation}
  \det \widetilde{\mathcal{K}} = 
    \det \widehat{\mathcal{K}} + 
    \det \widehat{\mathcal{K}}_{[\bm{F} \to n-row]}\; .
\end{equation}

It is now easy to see that \(\det \widehat{\mathcal{K}}\) is
proportional to the tau functions (Slavnov products), that is
\begin{equation}
    \det \widehat{\mathcal{K}}  = 
    \det(\widehat{\Omega})
    \det_{l,j}
    \left( \frac{z_l^{n-j}}{\prod_{s\neq l} (z_l - z_s)} \right)\\
    = \frac{\det(\widehat{\Omega})}{\Delta(\bm{z})}
\end{equation}
where \(\widehat{\Omega}\) is the matrix formed by the components
\(\widehat{\Omega}_{i,j}\). The above determinant is itself a tau function,
but we can also write is as
\begin{equation}
    \det \widehat{\mathcal{K}} = 
    \frac{\det(\Omega)}{\Delta(\bm{z})}
    \prod_{l=1}^n \left( 1 - \frac{1}{\lambda z_l} \right) 
    = \bm{\tau}(\bm{z}, \bm{v}) \prod_{l=1}^n \left( 1 - \frac{1}{\lambda z_l} \right) \; .
\end{equation}

Finally, using the elementary symmetric polynomials \(\sigma_p^{(n)}\), we write
\begin{equation}
  \prod_{l=1}^n \left( 1 - \frac{1}{\lambda z_l} \right) = 1 + 
  \sum_{p=1}^n (-1)^p \lambda^{-p} \sigma_p^{(n)}(\bm{z}^{-1})\; .
\end{equation}
where \( \bm{z}^{-1} = \{z_1^{-1}, z_2^{-1}, \dots , z_n^{-1}\}\).

Putting all these facts together, we write the Baker-Akhiezer function
as 
\begin{equation}
  \psi(\lambda; \bm{z}, \bm{v}) =
  e^{\xi(\bm{t}, \lambda)}  \left[ 1 + 
  \sum_{p=1}^n (-1)^p \lambda^{-p} \sigma_p^{(n)}(\bm{z}^{-1})
    + \frac{1}{\bm{\tau}(\bm{z}, \bm{v})}
    \det \widehat{\mathcal{K}}_{[\bm{F} \to n-row]}
  \right]\; .
\end{equation}
